Well-quasi-orderings on word languages

### 2025-09-08 src/amalgam.tex
todos: refactor + add subsections + add tightness

>>>>

\todo[inline]{%
  We have studied infinite words, but there are other ways to represent
  languages, such as finite automata, context-free grammars, or vector
  addition systems with states. In this section, we show that even for
  a generalisation of these systems, we can decide whether the language is
  well-quasi-ordered by the infix relation, and that in turns,
  they all fit in the bounded languages studied in the previous section.
}

### 2025-09-08 src/infixes.tex
fix: making llncs work
todos: refactor + add subsections + add tightness

\begin{theorem}[restate=bounded-language:thm,label=bounded-language:thm]
    %\label{bounded-language:thm}
    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
    $L$ is a \kl{well-quasi-order} when endowed with the 
    \kl{infix relation} if and only if it included in a finite union of 
    products $S_i \cdot P_i$ where 
    $S_i$ is a \kl{chain} for the \kl{suffix relation}, and 
    $P_i$ is a \kl{chain} for the \kl{prefix relation},
    for all $1 \leq i \leq n$.
\end{theorem}

@@

@@

\begin{corollary}
    \label{ordinal-invariants-bounded:cor}
    Let $L$ be a \kl{bounded language} of $\Sigma^*$
    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
    Then, the \kl{ordinal width} of $L$ is finite,
    its \kl{ordinal height} is at most $\omega$,
    and its \kl{maximal order type} is at most $\omega$.
\end{corollary}

@@

>>>>

\begin{theorem}
    \label{bounded-language:thm}
    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
    $L$ is a \kl{well-quasi-order} when endowed with the 
    \kl{infix relation} if and only if it included in a finite union of 
    products $S_i \cdot P_i$ where 
    $S_i$ is a \kl{chain} for the \kl{suffix relation}, and 
    $P_i$ is a \kl{chain} for the \kl{prefix relation},
    for all $1 \leq i \leq n$.
\end{theorem}

@@

\begin{corollary}
    \label{ordinal-invariants-bounded:cor}
    Let $L$ be a \kl{bounded language} of $\Sigma^*$
    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
    Then, the \kl{ordinal width} of $L$ less than $\omega^2$,
    its \kl{ordinal height} is at most $\omega$,
    and its \kl{maximal order type} less than $\omega^3$.
    Furthermore, those three bounds are tight.
\end{corollary}
\begin{proof}
  Upper bounds are a direct consequence of \cref{bounded-language:thm},
  and the tightness is witnessed by the 
  languages: 
  $L_k \defined \bigcup_{i = 2}^{k+1} (a b^i a)^* (b a^i b)^*$,
  that are \kl{bounded languages} of $\set{a,b}^*$,
  \kl{well-quasi-ordered} by the \kl{infix relation},
  and have \kl{ordinal width}, \kl{ordinal height} and
  \kl{maximal order type} respectively equal to $\omega \cdot k$, $\omega$ and $\omega^2 \cdot k$.
\end{proof}

\subsection{Downwards Closures and Bounded Languages}

@@

\begin{theorem}
  \label{bounded-wqo-dw-closed-regular:thm}
  Let $L$ be a \kl{downwards closed} and \kl{well-quasi-ordered} language
  of $\Sigma^*$ for the \kl{infix relation}. The following are equivalent:
  \begin{enumerate}
    \item $L$ is a \kl{bounded language},
    \item $L$ is a \kl{regular language},
    \item There exists a finite set $E \subset_{\mathsf{fin}} (\Sigma^*)^3$ such that
      $L = \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}$.
  \end{enumerate}
\end{theorem}

\subsection{General Downwards Closed Languages}

@@

\todo[inline]{aliaume: write it}
Let us conclude this section with a few remarks on the notion of
\kl{factor complexity} of languages.

\begin{lemma}
  There exists a language $L$ with exponential factor complexity 
  that is \kl{well-quasi-ordered} for the \kl{infix relation}.
\end{lemma}

\subsection{Decision Procedures}
